---
title: "Data Import"
author: "Jana Lee"
date: "9/17/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Load in a dataset for from FAS_litters
Always import the relative path. Never the absolute path!
```{r}
litters_data = read_csv(file = "./data_import_examples/FAS_litters.csv")
litters_data = janitor::clean_names(litters_data)
# Janitor::clean_names: Will take whatever the column name is and convert this to lower snake case.
```

## Learning Assessment 1, load in pups data
```{r}
pups_data = read_csv(file = "./data_import_examples/FAS_pups.csv")
pups_data = janitor::clean_names(pups_data)
```

```{r}
tail(litters_data, 5)
#Returns the last parts of a vector, matrix table, data frame, or function. Use head, str, tail.

skimr::skim(litters_data)
#Skimr is another way to look through the data.

view(litters_data)
# This will open a new tab for us to view the data.
```

## Arguments in the read statement
If data is clean, you can use col_names = true
```{r}
litters_data = read_csv(file = "./data_import_examples/FAS_litters.csv",
  skip = 10, col_names = FALSE)
```


##parsing columns - back ticks show that the column name is NOT in a standard format. It's a warning to the code!
```{r}
litters_data = read_csv(file = "./data_import_examples/FAS_litters.csv",
col_types = cols(
    Group = col_character(),
    `Litter Number` = col_character(),
    `GD0 weight` = col_double(),
    `GD18 weight` = col_double(),
    `GD of Birth` = col_integer(),
    `Pups born alive` = col_integer(),
    `Pups dead @ birth` = col_integer(),
    `Pups survive` = col_integer()
  )
)
```

## Learning assessment 2, parsing example
```{r}
pups_data = read_csv("./data_import_examples/FAS_pups.csv", col_types = "ciiiii")
skimr::skim(pups_data)
```


## Read in an Excel file...
```{r}
library(readxl)
mlb11_data = read_excel(path = "./data_import_examples/mlb11.xlsx")
```


## Read in SAS...
```{r}
pulse_data = haven::read_sas("./data_import_examples/public_pulse_data.sas7bdat")
```

## Learning Assessment 3, Importing
```{r}
pups_base = read.csv("./data_import_examples/FAS_pups.csv")

pups_readr = read_csv("./data_import_examples/FAS_pups.csv")

pups_base

pups_readr

# read_csv produces tibbles which are similar to the base R data frames produced by read. csv

```


## Exporting data
```{r}
library(readxl)
mlb11_data = read_excel(path = "./data_import_examples/mlb11.xlsx")
# Load in the dataset

mlb11_data_subset = 
  read_excel(
    path = "./data_import_examples/mlb11.xlsx",
    range = "A1:D7")
# Make a new sub dataset.

write_csv(mlb11_data_subset, path = "./data_import_examples/mlb_subset.csv")
# Use the write function
```


