---
title: "Data Manipulation"
author: "Jana Lee"
date: "9/19/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Import datasets litters & pups

```{r, import}
litters_data = read_csv("./data_import_examples/FAS_litters.csv")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv( "./data_import_examples/FAS_pups.csv")
pups_data = janitor::clean_names(pups_data)
```

## Select
Use when you only need a subset of the columns in the data.
Extract only what you need can de-cluter the data. This is strictly for column
names.
```{r}
select(litters_data, group, litter_number)

select(litters_data, group, litter_number, gd0_weight)

select(litters_data, group, litter_number, gd0_weight, everything())

select(litters_data, -group)

select(litters_data, group, litter_number, gd0_weight, pups_born_alive)

# Renamed group function
select(litters_data, GROUP = group, litter_number)

# renaming the variables that we have 
rename(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)

```

## Filtering !!!
Filtering for the Con7 group. You can check this by looking at the tiblble -> 
7*8
```{r, filter}
filter(litters_data, group == "Mod8")
# double = tries to see if those values are equal to each other.

filter(litters_data, gd_of_birth == 20)

filter(litters_data, gd_of_birth < 20)

filter(litters_data, gd_of_birth >= 20)

#first we want to filter by pups_born_alive, then we want Con7.
filter(litters_data, pups_born_alive < 6, group == "Con7")

# can put ! before any data statement. It mean NOT something. so opposite of what we might want

# group %in% c("Con7", "Con8") --> can filter for everything in between

# how to do OR statement?
filter(litters_data, group %in% c("Con7", "Mod8"))

# DON'T DO THIS:
# filter(litters_data, is.na(gd0_weight))

# This one is better. Drop_Na is better
drop_na(litters_data, gd0_weight)

```

## Mutate
Changing them or creating new ones with mutate.
```{r}
mutate(litters_data,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
```

## Arrange
Maybe, I want this to be in some order that it's not currently in.
```{r}
arrange(litters_data, pups_born_alive)

# This will arrange pup born alive and then weight
arrange(litters_data, pups_born_alive, gd0_weight)

```





