---
title: "Tidy_Data"
author: "Jana Lee"
date: "9/24/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library (tidyverse)
```

## Wide to long

```{r}
pulse_data =
  haven::read_sas("./data_import_examples/public_pulse_data.sas7bdat") %>%
  janitor::clean_names()
```

BDI score is spread across 4 columns, corresponding to 4 observation times. We can fix the problem with pivot_longer.

We want a visit ID number that removes the bdi_score... etc. We can use the names_prefix to get RID of this.
```{r}
pulse_tidy_data = 
  pivot_longer(
    pulse_data, 
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
) %>%
  mutate(
    visit = recode(visit, "bl" = "00m"),
  )
```

## separate in litters
```{r}
litters_data = 
  read_csv("./data_import_examples/FAS_litters.csv", col_types = "ccddiiii") %>% 
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  mutate(
    dose = str_to_lower(dose),
    wt_gain = gd18_weight - gd0_weight) %>%
  arrange(litter_number)
```

## pivot_wider
Deliberately untidying this
```{r}
analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
)

pivot_wider(
  analysis_result,
  names_from = "time",
  values_from = "mean",
)
```

What happens when we have data spread across multiple tables?
```{r}
fellowship_data = 
  readxl::read_excel("./data_import_examples/LotR_Words.xlsx", range = "B3:D6") %>%
  mutate(movie = "fellowship_ring")

two_towers = 
  readxl::read_excel("./data_import_examples/LotR_Words.xlsx", range = "F3:H6") %>%
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data_import_examples/LotR_Words.xlsx", range = "J3:L6") %>%
  mutate(movie = "return_king")
```

## bind_rows
```{r}
lotr_tidy = 
  bind_rows(fellowship_data, two_towers, return_king) %>%
  janitor::clean_names() %>%
  pivot_longer(
    female:male,
    names_to = "sex", 
    values_to = "words") %>%
  mutate(race = str_to_lower(race)) %>% 
  select(movie, everything()) 
```


## Joining datasets
Adding the pup_data and litter_data first. Then, making fas_data (new dataset) to join the pups data together. 
```{r}
pup_data = 
  read_csv("./data_import_examples/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("./data_import_examples/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group))

# We are specifying that we start with pup_data first and then add the litter_data)
# if you don't tell the left_join function with by, it might get you into trouble. It is smart and tries to join by itself
fas_data =
  left_join(pup_data, litter_data, by = "litter_number")

# Joining by multiple: ... by = c(...)

```

